---
title: "AE 07: Partitioning variability"
subtitle: "Price of textbooks"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```

## Announcements

- Ed Discussion for today's lecture: https://edstem.org/us/courses/8350/discussion/606083

- Prep Quiz 02 **due Wed, Sep 15 at 3:30p**

- HW 01 **due Wed, Sep 15 at 11:59p**

## Question about HW? 

## Questions from prep assignment? 


##  AE 07: Price of textbooks

- Go to the ae-07-[GITHUB USERNAME] repo in the GitHub course organization: https://github.com/sta210-fa21

- Access the RStudio Docker containers at https://vm-manage.oit.duke.edu/containers

- See the [Lab 01 instructions](https://sta210-fa21.netlify.app/labs/lab-01-ikea.html) for details on cloning the repo and starting a new project in RStudio.

## HW 01
