---
title: "AE 06: Prediction for SLR"
subtitle: "Popular songs on Spotify"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```

## Announcements

- Ed Discussion for today's lecture: https://edstem.org/us/courses/8350/discussion/594810

- HW 01 **due Wed, Sep 15 at 11:59p**

- Look out for email about this week's lab

## Workflow tips

- Save your work often!!
- Knit, commit, and push your work often!!
- RStudio containers are periodically refreshed, so any unsaved work will be lost (e.g., work you do in the console)
  - Any saved files will not be affected
- If you're having trouble with the dockers submit this form: https://duke.qualtrics.com/jfe/form/SV_6s3DOsiCIzNya9w

## Questions from prep assignment? 

##  AE 06: Popular songs on Spotify

- Go to the ae-06-[GITHUB USERNAME] repo in the GitHub course organization: https://github.com/sta210-fa21

- Access the RStudio Docker containers at https://vm-manage.oit.duke.edu/containers

- See the [Lab 01 instructions](https://sta210-fa21.netlify.app/labs/lab-01-ikea.html) for details on cloning the repo and starting a new project in RStudio.

## HW 01
