---
title: "STA 210 Lab 04"
date: "09.30.2021"
output:
  xaringan::moon_reader:
    css: "sta210-slides.css"
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

## Today's lab 

- ANOVA app activity 

- Lab 04: ANOVA with Palmer Penguins

---

## ANOVA app activity (~ 5 - 10 minutes)

For this activity: 

- Complete the ANOVA App Activity assessment on Sakai. It includes questions about ANOVA and reflection questions. This portion will be graded for completion and will be part of the Lab 04 grade. 
  - We encourage you to make an honest attempt at each question, as it can help you gauge your understanding of ANOVA. 
  
---

## Team Workflow reminders

Once you have completed the ANOVA activity, you will work on Lab 04 with your lab team. A few workflow reminders: 

- Only one team member should type at a time to reduce concflicts in your RMarkdown and PDF documents.

  - Every team member should still be engaged in discussion for all questions, even if it's not your turn type. 
  
- Don't forget to **pull** to get your teammates' updates before making changes to the .Rmd file 

- Only **one** submission for the team on Gradescope. Select every team members' name in Gradescope. 
  
  
---

class: middle, center

## Please raise your hand if you need help as you work on the lab.
  

