---
title: "AE 04: Inference for simple linear regression"
subtitle: "Bike rentals in Capital Bikeshare"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```

## Announcements

- Thursday's labs will all be online. Find the Zoom link for your lab section under the  **Zoom Meetings** tab in Sakai. 
- Office hours started this week. Schedule available on the [syllabus](https://sta210-fa21.netlify.app/syllabus/).

- Complete [Get to Know You survey](https://duke.qualtrics.com/jfe/form/SV_eyUiuqZKsJaqHWK) by **today**

- [Week 03](https://sta210-fa21.netlify.app/prepare/week-03.html) prepare
  - Prep Quiz 01 due Wed, Sep 08 at 3:30p

- Lab 01 accepted until Sep 02 at 11:59p with no late penalty


## Questions about logistics? 

## Questions from prepare assignment?

##  AE 04: Inference for simple linear regression

- Go to the ae-04-[GITHUB USERNAME] repo in the GitHub course organization: https://github.com/sta210-fa21

- Access the RStudio Docker containers at https://vm-manage.oit.duke.edu/containers

- See the [Lab 01 instructions](https://sta210-fa21.netlify.app/labs/lab-01-ikea.html) for details on cloning the repo and starting a new project in RStudio.

