---
title: "AE 05: Conditions for simple linear regression"
subtitle: "Popular songs on Spotify"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```

## Announcements

- Lab 02 due today at 11:59p

- Prep Quiz 01 due Wed, Sep 08 at 3:30p

- Ed Discussion for today's lecture: https://edstem.org/us/courses/8350/discussion/590317


## Code chunk names

- Why we name code chunks
- Don’t include spaces in code chunk names!
  - `r emo::ji("frowning_face")` {r fit model}
  - `r emo::ji("slightly_smiling_face")` {r fit-model}

## Questions from prep assignment? 

##  AE 05: Popular songs on Spotify

- Go to the ae-05-[GITHUB USERNAME] repo in the GitHub course organization: https://github.com/sta210-fa21

- Access the RStudio Docker containers at https://vm-manage.oit.duke.edu/containers

- See the [Lab 01 instructions](https://sta210-fa21.netlify.app/labs/lab-01-ikea.html) for details on cloning the repo and starting a new project in RStudio.

