---
title: "AE 09: Analysis of variance"
subtitle: "Duke Lemurs"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_height: 3
    fig_width: 6
    theme: readable
---

```{r setup, echo = F}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center")
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```

## Announcements

- Ed Discussion for today's lecture: https://edstem.org/us/courses/8350/discussion/627393

- Lab 03 **due today at 11:59p**

- HW 02 **due Wednesday, Sep 22 at 11:59p** 

- Exam 01 open *Thursday, Sep 23 at 9a - Friday, Sep 24 at 11:59p*
  - No lab Thursday Sep 23

- Exam 01 review Wednesday 
  - [Click here](https://forms.gle/4bJbKSUWtNn8E18t8) to submit questions. 

## Question about HW or lab? 

##  AE 09: ANOVA with Duke lemurs

- Go to the ae-09-[GITHUB USERNAME] repo in the GitHub course organization: https://github.com/sta210-fa21

- Access the RStudio Docker containers at https://vm-manage.oit.duke.edu/containers

- See the [Lab 01 instructions](https://sta210-fa21.netlify.app/labs/lab-01-ikea.html) for details on cloning the repo and starting a new project in RStudio.

