---
title: "STA Lab"
subtitle: "09.09.21"
author: ""
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta210-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---

```{r child = "setup.Rmd"}
```

## Set up 

- Today's lab will focus on working on HW 01

- The data include county-level demographic and county-level information and variables to help with creating maps 

- The demographic and election data are from the Economist and were used in the July 2021 article [" In-person voting really did accelerate covid-19’s spread in America"](https://www-economist-com.proxy.lib.duke.edu/graphic-detail/2021/07/10/in-person-voting-really-did-accelerate-covid-19s-spread-in-america)

---

## Goals

- Use regression to describe the relationship between a county's political leanings (measured by % of votes for Trump in 2016) and the percent of votes cast in-person in the 2020 election for counties in North Carolina. 
- Use maps to conduct exploratory data analysis and check the model conditions 
- Use a random sample of of US counties to draw conclusions about the relationship between political leanings and percent of votes cast in-person in the 2020 election
  
---

## Exercise 1

- What summary statistics best represent the center and spread of the distribution? why? 

---

## Exercise 2 - the code

- **Goal**: We will look at the response variable by county. 

- `county_map_data`: contains the latitude, longitude for the counties in NC

- What does the code below do? 

```{r eval = F, echo = T}
election_map_data <- left_join(election_nc, county_map_data)
```

---

## Exercise 2 - the code

[`geom_polygon`](https://ggplot2.tidyverse.org/reference/geom_polygon.html): Function use to plot polygons. The coordinates for each vertex is on a separate row. The vertices for a given `group` are connected and filled in.

.small[
```{r eval = F, echo = T}
ggplot() + 
  geom_polygon(county_map_data, 
               # make a map of NC 
               mapping = aes(x = long, y = lat, group = group),
               # fill each polygon (county) light gray & 
               # outline in white
               fill = "lightgray", color = "white") + 
```
]
---

## Exercise 2 - the code

[`geom_polygon`](https://ggplot2.tidyverse.org/reference/geom_polygon.html): Function use to plot polygons. The coordinates for each vertex is on a separate row. The vertices for a given `group` are connected and filled in.

.small[
```{r eval = F, echo = T}
ggplot() + 
  geom_polygon(county_map_data, 
               # make a map of NC 
               mapping = aes(x = long, y = lat, group = group),
               # fill each polygon (county) light gray & 
               # outline in white
               fill = "lightgray", color = "white") + 
  geom_polygon(election_map_data, 
               mapping = aes(x = long, y = lat, group = group, 
               # fill each polygon (county) based on inperson_pct 
               fill = inperson_pct)) + 
  
```
]

---

## Exercise 2 - the code

.small[
```{r eval = F, echo = T}
ggplot() + 
  geom_polygon(county_map_data, 
               mapping = aes(x = long, y = lat, group = group),
               fill = "lightgray", color = "white") + 
  geom_polygon(election_map_data, 
               mapping = aes(x = long, y = lat, group = group, 
               fill = inperson_pct)) + 
  # add labels
  labs(x = "_____", 
       y = "_____", 
       fill = "_____", 
       title = "_____") +
  # use viridis color palette
  scale_fill_viridis() 
```
]

Read more about the [viridis color palettes](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html). 
  
---

## Exercise 2 

- If there is no association between county and percent of in-person voting, what would you expect the color pattern to look like on the map? 

- Which plot do you prefer to help understand the response variable - the histogram or the map? Why? 