---
title: "STA 210 Lab"
subtitle: "09.16.21"
author: ""
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "sta210-slides.css"
    logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---

```{r child = "setup.Rmd"}
```

## Goals

- Meet your teammates!

- Ice breaker / get to know your team 

- Team agreement

- Lab 03: Coffee ratings

---

## Meet your Teammates!

- [Click here](https://prodduke-my.sharepoint.com/:x:/g/personal/mt324_duke_edu/EcxaeUdU8mRIiJHzEee-UBgBDpLfk1uNAgsnoEe4L2EdjQ?e=X4ndeQ) to find your team number.

- Sit with your team. 


---

## Icebreaker (5 minutes)

- Quick introductions: Name and hometown

- Choose a reporter: The person whose birthday is closest to September 16

- Identify 8 things everyone in the group has in common:    
  - Not clothes (e.g. we're all wearing shoes)! 
  - Not body parts, (e.g. we all have a nose.)!

```{r echo = FALSE}
library(countdown)
countdown(minutes = 5, seconds = 0)
```

---

## Team name + agreement

- Come up with a team name. You can't have the same name as another group in the class, so be creative! 
  - Your TA will get your team name by the end of lab.

- Fill out the team agreement. The goals of the agreement are to...
  - Gain a common understanding of the team's goals and expectations for collaboration
  - Make a plan for team communication
  - Make a plan for working outside of lab. 
  

---

## Demo: 

- Collaboration using a single GitHub repo
- **Pull** work from GitHub into RStudio

---

## Team workflow 

- Only one team member should type at a time. There are markers in today's lab to help you determine who's turn it is to type. 

  - Every team member should still be engaged in discussion for all questions, even if it's not your turn type. 
  
- Don't forget to **pull** to get your teammates' updates before making changes to the .Rmd file 

- Only **one** submission for the team on Gradescope.

---

## Tips for working on a team

- Do not pressure each other to finish early; use the time wisely to really learn the material and produce a quality report.

- The labs are structured to help you learn the steps of a data analysis. Do not split up the lab among the team members; work on it together in its entirety.

- Everyone has something to contribute! Use the lab groups as an opportunity to share ideas and learn from each other.




