---
title: "Lab 02: Simple linear regression"
subtitle: "due Mon, September 06 at 11:59p ET"
output: 
  tufte::tufte_html:
    css: "sta210-labs.css"
    tufte_variant: "envisioned"
    highlight: pygments
    fig_height: 4
    fig_width: 6
    toc: true
    toc_depth: 1
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE)
```


# Introduction

```{marginfigure}
R is the name of the programming language itself and RStudio is a convenient interface.
```

This lab will go through much of the same workflow we've demonstrated in class.
The main goal is to reinforce our demo of R and RStudio, which we will be using throughout the course both to learn the statistical concepts discussed in the course and to analyze real data and come to informed conclusions.

```{marginfigure}
git is a version control system (like “Track Changes” features from Microsoft Word but more powerful) and GitHub is the home for your Git-based projects on the internet (like DropBox but much better).
```

An additional goal is to reinforce git and GitHub, the collaboration and version control system that we will be using throughout the course.

As the labs progress, you are encouraged to explore beyond what the labs dictate; a willingness to experiment will make you a much better programmer.
Before we get to that stage, however, you need to build some basic fluency in R.
Today we begin with the fundamental building blocks of R and RStudio: the interface, reading in data, and basic commands.

To make versioning simpler, this and the next lab are solo labs. In the future, you'll learn about collaborating on GitHub and producing a single lab report for your lab team, but for now, concentrate on getting the basics down.

## Learning goals 

By the end of the lab, you will...

- Be familiar with the workflow using RStudio and GitHub
- Gain practice writing a reproducible report using RMarkdown
- Practice version control using GitHub## Getting started 

# Getting started 

-   Go to the [**sta210-fa21**](https://www.github.com/sta210-fa21) organization on GitHub. Click on the repo with the prefix **lab-02**. It contains the starter documents you need to complete the lab.

- Clone the repo and start a new project in RStudio. See the [Lab 01 instructions](https://sta210-fa21.netlify.app/labs/lab-01-ikea.html) for details on cloning a repo, starting a new R project and configuring git. 

## Packages

We will use the following packages in today’s lab.

```{r eval = F}
library(tidyverse)
library(broom)
library(knitr)
```

## The Data 


The data for this lab is from the `scorecard` data set in the **rcfss** R package. It includes information originally obtained from the [U.S. Department of Education’s College Scorecard](https://collegescorecard.ed.gov/data/) for 1753 colleges and universities during the 2018 - 2019 academic year. 

We will use the following variables in this lab: 

- `admrate`: Undergraduate admissions rate (from 0-100%)
- `cost`: The average annual total cost of attendance, including tuition and fees, books and supplies, and living expenses
- `type`: Type of college (Public; Private, nonprofit; Private, for-profit)

[Click here](https://uc-cfss.github.io/rcfss/reference/scorecard.html) to see a full list of variables and definitions. 

Use the code below to load the data set. 

```{r eval = F}
scorecard <- read_csv("data/scorecard.csv")
```


# Exercises

1. Create a histogram to examine the distribution of `admrate` and calculate summary statistics for the center (mean and median) and the spread (standard deviation and IQR). 

2. Use the results from the previous exercise to describe the distribution of `admrate`. Include the shape, center, spread, and if there are potential outliers. 

3.  Plot the distribution of `cost` and calculate the appropriate summary statistics. Describe the distribution of `cost` (shape, center, and spread, and outliers) using the plot and appropriate summary statistics.


4. The goal of this analysis is to fit a regression model that can be used to understand the variability in the cost of college based on the admission rate. Before fitting the model, let's look at the relationship between the two variables. Create a scatterplot to display the relationship between cost and admissions rate. Use the scatterplot to describe the relationship between the two variables.


5. Does the relationship between cost and admissions rate differ by type of college? Modify the plot from the previous exercise to explore the relationship by type of college. Describe two new observations from this scatterplot that you didn’t see in the previous plot.

6. Describe two new observations from the scatterplot in Exercise 5 that you didn't see in the scatterplot from Exercise 4. 


7. Use the lm() function to fit a simple linear regression model to describe the relationship between a college’s admission rate and cost. Use the code below to assign your model a name, and use the tidy() and kable() functions to neatly display the model output. Replace X and Y with the appropriate variable names.

```{r eval = F}
_____ <- lm(Y ~ X, data = _____)

tidy(_____) %>% # output model

  kable(format = "markdown", digits = 3) # format model output
```

8. Let's interpret the coefficients of the model from the previous exercise. 
  - Interpret the slope in the context of the problem.
  - Does the intercept have a meaningful interpretation? If so, write the interpretation in the context of the problem. Otherwise, explain why the interpretation is not meaningful.

9. Is the linear relationship between cost and admissions rate and statistically significant? In your response

- State the null and alternative hypotheses used to answer this question in words and in mathematical notation.

- What is the test statistic? Write the equation used to calculate the test statistic.

- What distribution was used to calculate the p-value?

- State your conclusion for the test in the context of the data.

10. Interpret the 95% confidence interval for the slope in the context of the data.

# Submission

In this class, we'll be submitting PDF documents to Gradescope.

**Before you wrap up the assignment, make sure all documents are updated on your GitHub repo. We will be checking these to make sure you have been practicing how to commit and push changes.**

**Remember -- you must turn in a PDF file to the Gradescope page before the submission deadline for full credit.** 

To submit your assignment: 

- Go to http://www.gradescope.com and click *Log in* in the top right corner. 

- Click *School Credentials* `r emo::ji("arrow_right")` *Duke NetID* and log in using your NetID credentials.

- Click on your *STA 210* course. 

- Click on the assignment, and you'll be prompted to submit it.

- Mark the pages associated with each exercise. All of the pages of your lab should be associated with at least one question (i.e., should be "checked"). 

- Select the first page of your .PDF submission to be associated with the *"Workflow & formatting"* section.

# Grading (50 pts)

<br>

| Component | Points |
|:----------|:-------|
| Ex 1 - 10     | 45      |
| Workflow & formatting     | 5    

**Grading notes**:

-   The "Workflow & formatting" grade is to assess the reproducible workflow. This includes having at least 3 informative commit messages and updating the name and date in the YAML.
